---
title: "OHI: Red Sea and Gulf of Aden"
format: html
editor: visual
---

```{r}
#| echo: false
#| warning: false
#| message: false
library(sf)
library(tidyverse)
library(here)
library(leaflet)
library(rnaturalearth)
```

## Overview of region

The Red Sea and Gulf of Aden are unique biological areas that provide many services that humans rely upon.

The Red Sea is one of the most saline bodies of water on Earth, which has fostered the evolution of a unique and diverse ecosystem. The region is home to a variety of marine species, including over 1,200 species of fish, of which about 10% are found no where else. The coral reefs in the Red Sea are also among the most diverse in the world.

Similar to the Red Sea, the Gulf of Aden is home to a wide variety of marine life, although its ecological diversity is less studied. The mangroves and coastal wetlands around the Gulf of Aden are crucial for biodiversity, providing breeding grounds for many fish species and habitat for a variety of wildlife.

### Physical description

The Red Sea is a seawater inlet of the Indian Ocean, lying between Africa and Asia. It is known for its remarkable depth and narrow width.

The waters of Red Sea are mostly divided between the countries of Saudi Arabia, Egypt, Sudan, and Eritrea. However, Jordan and Israel have small EEZs in the northern region, located in the Gulf of Aqaba. And, Djibouti, located at the southern extremity, has a smaller EEZ.

There are two disputed EEZ regions located in Red Sea. One off of the Halaib Triangle between Sudan and Egypt; and the Doumeira Isands between Eritrea and Djibouti.

The southern tip of the Red Sea opens into the Gulf of Aden via the Bab el Mandeb strait. The Gulf of Aden is shared among three nations, Yemen on the Arabian Peninsula and Djibouti and Somalia in Africa.

The Exclusive Economic Zone (EEZ) of Somalia is complex due to issues of sovereignty and international recognition between the Federal Republic of Somalia and the self-declared Republic of Somaliland. Although Somaliland declared independence from Somalia in 1991, it is not internationally recognized as a sovereign state. As a result, in terms of international law and most international agreements, the waters off the coast of Somaliland are considered part of Somalia's EEZ.

In practice, the government of Somaliland administers and lays claim to the waters off its coast as part of its self-declared sovereignty. However, these claims are not recognized by the international community, and any agreements or exploration licenses issued by Somaliland in these waters are not acknowledged by most countries and international organizations.

```{r}
#| echo: false
#| warning: false
#| message: false

eez <- st_read(here("_spatial/data/eez.shp"), quiet=TRUE) %>%
  arrange(country)
#eez$wtr_bdy
#eez$country
eez$color <- c("#FB2500", "#FB2500", "#1DA08A","#F2AD01","darkgray", "#F98400","#B13259", "#E2D200", "#7B1C6B", "#5ABCD6","darkgray",  "#F8AFA8", "#F8AFA8")

red_sea <- st_read(here("_spatial/data/water_bodies.shp"), quiet=TRUE) %>%
  filter(name== "Red Sea")
gulf_aden <- st_read(here("_spatial/data/water_bodies.shp"), quiet=TRUE) %>%
  filter(name== "Gulf of Aden")

land <- ne_countries(scale = "medium", returnclass = "sf") %>%
  filter(sov_a3 %in% c("DJI", "EGY", "ERI", "SOM", "SOL", "ISR", "JOR" , "SAU", "SDN", "YEM"))

# leaf maps: https://leaflet-extras.github.io/leaflet-providers/preview/index.html
#st_bbox(eez)

map <-  leaflet() %>%
  addProviderTiles("CartoDB.Positron") %>%
    addPolygons(data = red_sea, 
              fillColor = NA, 
              weight = 2, 
              color = "brown", 
              opacity = 1) %>%
    addPolygons(data = gulf_aden, 
              fillColor = NA, 
              weight = 2, 
              color = "#5ABCD6", 
              opacity = 1) %>%
  addPolygons(data = eez,
              fillColor = ~color,
              weight = 0, 
              fillOpacity = 0.5, 
              color=NA,
              popup = ~paste(dscrptn, "<br>",
                             "km<sup>2</sup> = ", are_km2)) %>%
  addPolygons(data = land,
              fillColor = "lightgray",
              weight = 2, 
              fillOpacity = 0.5, 
              color="darkgray",
              popup = ~paste(formal_en)) %>%
  setView(lng=41, lat=20, zoom = 5)

map


```

### Challenges

Both the Red Sea and Gulf of Aden face environmental challenges from overfishing; pollution from oil extraction, shipping, and land-based sources; and climate change.

The Gulf of Aden also has a history of piracy, and the lack of effective governance and security in the waters has led to illegal fishing and dumping.

International cooperation is necessary to mitigate the environmental challenges facing the region. Initiatives could include establishing marine protected areas, enforcing stricter fishing regulations, and implementing measures to reduce pollution and tackle the effects of climate change. Additionally, regional cooperation and governance are crucial in addressing issues like piracy and illegal fishing in the Gulf of Aden.
