---
title: "Ocean Health Index"
subtitle: "Red Sea and Gulf of Aden"
---

```{r}
#| echo: false
#| warning: false
#| message: false

source("_data/functions/start_script.R")
source("_data/functions/score_display_functions.R")
score <- score(x="Index", ohi_scores)

```

:::: {.columns} 

::: {.column width="50%"}

<br>
<br>
Using data from the global 2023 Ocean Health Index analysis, the overall score for the Red Sea and Gulf of Aden region was `r score`.
:::

::: {.column width="50%"}

```{r, fig.width=10, fig.height=3}
#| echo: false
#| warning: false
#| message: false

dial(score,colors)

```

:::

::::

The interactive map below shows how different countries contributed to this score.

```{r, fig.width=3, fig.height=1.5}
#| echo: false
#| warning: false
#| message: false
map_figure(ohi_scores=ohi_scores, goal="Index")

```

<br>

Each region's score is the average of 10 goals that describe all the ways human value and rely on marine systems. 

You can navigate through the menu at the right to see each goal score and learn more about each goal.

:::{.callout-important}
## Regional models will produce more relevant OHI scores!

This score is the EEZ weighted average score from the OHI global assessment. A regional assessment can be better for several reasons: 

- the global analysis requires using data the most complete global data, however,  local data may be available that is better!

- the goal models can be adapted to better capture the values and concerns of the region. 

- the goals can be weighted differently, and can even be changed.

:::
