---
title: ""
format: html
---
# ![](figs/artisanal-fishing-opportunities.svg){width=8% height=8% color="red"} Artisanal fishing opportunities 

**This goal measures whether people have the opportunity to fish at a small local scales.**

## Overview

Artisanal fishing, often called small-scale fishing, provides a critical source of food, nutrition, poverty alleviation, and livelihood opportunities for many people around the world, especially in developing nations. Small-scale fishing is also an important component of cultural identity and heritage.

## Scores
```{r}
#| echo: false
#| warning: false
#| message: false

source("../_data/functions/start_script.R")
#source("_data/functions/start_script.R")
source("../_data/functions/score_display_functions.R")
#source("_data/functions/score_display_functions.R")
score <- score(x="AO", ohi_scores)

```

:::: {.columns} 

::: {.column width="50%"}

<br>

**The Artisanal Fishing Opportunities score for the Red Sea and Gulf of Aden region was `r score`.**

<span style="color: gray;">Overall, these scores are fairly high which indicates that demand for artisanal fishing is being met using lawful and sustainable methods, or alternatively, there is little economic need for artisanal fishing (e.g., Saudi Arabia, Israel, Yemen). The lowest scoring regions are Eritrea (high need, fairly low access) and Sudan (high need, small scale fisheries could be better managed).</span>
:::

::: {.column width="50%"}

```{r, fig.width=10, fig.height=3}
#| echo: false
#| warning: false
#| message: false

dial(score,colors)

```

:::

::::

The interactive map below shows the scores for the different countries.

```{r, fig.width=3, fig.height=1.5}
#| echo: false
#| warning: false
#| message: false
map_figure(ohi_scores=ohi_scores, goal="AO")

```

<br>


## Model description
This goal measures how easily people can access ocean resources when they *need* them and whether the stocks being harvested are sustainably managed. 

The model has three components:

-   *Need:* The percent of population below the poverty line. The global OHI model uses per capita GDP as a proxy for need.

-   *Access:* Institutional permits and regulations encouraging small fisheries and providing access to coastal regions. Data is from the UN sustainable development goal 14.b.1 which measures the degree of application of a legal, regulatory, policy, or institutional framework which recognizes and protects access rights for small-scale fisheries.

- *Sustainability:* Sustainability of fishing methods and/or the health of targeted species. Sustainable gears are traditional hook and line, as opposed to unsustainable practices such as dynamite and gill net fishing. The global assessment uses the same stock status data used to calculate the fisheries subgoal scores.

A high score indicates that the demand for artisanal fishing is being met using lawful and sustainable methods, or alternatively, there is little economic need for artisanal fishing. A low score indicates that regions are not achieving or allowing sustainable artisanal fishing opportunities to be realized.

## A closer look at the data

The biggest factor driving scores, by far, is the current status component. Here, we take a closer look at the data underlying the status scores for each country with EEZ territory in the Red Sea and Gulf of Aden.

### Current status
```{r}
#| echo: false
#| warning: false
#| message: false

goal <- "ao"
rsga_rgns <- read_csv(here("_data/output/rsga_ohi_rgns.csv"))

# list of OHI datafiles:
datafiles <- read_csv("https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/metadata_documentation/layers_eez_base.csv") %>%
  filter(str_detect(layer, goal)) 

# prep files:
prep <- read_csv("https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/metadata_documentation/layers_eez_dataprep.csv")

```

#### Economic need for small scale marine fisheries 

As a proxy for subsistence fishing need, we use the inverse of per capita purchasing power parity (PPP) adjusted gross domestic product (GDP), or GDPpcPPP. These data are rescaled to values between 0 and 1 by taking the natural log of the data and dividing all the values by the GDPpcPPP value that corresponds to the 95th quantile across all years (from 2005 to most recent data) and countries. 

Saudi Arabia, Israel, and Yemen have low economic need, whereas, the other countries have relatively high economic mean.

For more details see the data preparation [script](https://ohi-science.org/ohiprep_v2023/globalprep/ao/v2023/ao_need_data_prep.html).

::: {.callout-note collapse=true}

## Need: Country scores (click to expand)
```{r}
#| echo: false
#| warning: false
#| message: false


ao_need <- filter(datafiles, layer=="ao_need") 

ao_need_df  <- read_csv(file.path("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2023/gh-pages/",
gsub("ohiprep:", "", ao_need$dir), ao_need$fn)) 

ao_need_df <- left_join(rsga_rgns, ao_need_df, by="rgn_id") %>%
  group_by(iso3c) %>%
  filter(year==max(year)) %>%
  select(country = region_name, `economic need` = value) %>%
  mutate(`economic need` = round((1 - `economic need`), 2)) %>%
  arrange(`economic need`)

  datatable(ao_need_df,
          options = list(pageLength = 10, autoWidth = TRUE),
          filter = 'top')
```

:::

#### Access to small scale marine fisheries 
These data are from the [UN sustainable development goal 14.b.1](https://www.fao.org/sustainable-development-goals/indicators/14b1/en/), “Degree of application of a legal/regulatory/policy/institutional framework which recognizes and protects access rights for small-scale fisheries (level of implementation: 1 lowest to 5 highest)”. We linearly rescale these data to have values between 0 and 1.

All Red Sea and Gulf of Aden regions have high access scores from 0.8-1, except Eritrea which is 0.6.

::: {.callout-note collapse=true}

## Access to artisanal fisheries: Country scores (click to expand)

```{r}
#| echo: false
#| warning: false
#| message: false

ao_access <- filter(datafiles, layer=="ao_access") 

ao_access_df  <- read_csv(file.path("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2023/gh-pages/",
gsub("ohiprep:", "", ao_access$dir), ao_access$fn)) 

ao_access_df <- left_join(rsga_rgns, ao_access_df, by="rgn_id") %>%
  group_by(iso3c) %>%
  filter(year==max(year)) %>%
  select(country = region_name, `access` = value) %>%
  arrange(`access`)

  datatable(ao_access_df,
          options = list(pageLength = 10, autoWidth = TRUE),
          filter = 'top')
```

:::


#### Sustainable harvest of artisanally fished stocks 

This is calculated using the same data as the fisheries subgoal, but using only the artisanal catch. For each region, we average the stock status scores weighted by the total catch. The country level catch data (tonnes per year) is from FAO and the B/Bmsy (the biomass of the population relative to the biomass at maximum sustainable yield) are from RAM when available. Otherwise, we use the CMSY approach to estimate B/Bmsy. 

The scores are relatively high for all the Red Sea and Gulf of Aden regions. They are around 0.8 for for Sudan, Egypt, and Israel; and > 0.9 for the remaining countries.

For more information about the methods, check out the scripts used to prepare these data: ([preparing catch](https://ohi-science.org/ohiprep_v2023/globalprep/ao/v2023/ao_catch_prep_saup.html) and [estimating score]( 
https://ohi-science.org/ohiprep_v2023/globalprep/ao/v2023/ao_stock_status_saup.html)

::: {.callout-note collapse=true}

## Sustainable harvest of artisanal stocks: Country scores (click to expand)

```{r}
#| echo: false
#| warning: false
#| message: false


ao_sust <- filter(datafiles, layer=="ao_sust") 

ao_sust_df  <- read_csv(file.path("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2023/gh-pages/",
gsub("ohiprep:", "", ao_sust$dir), ao_sust$fn)) 

ao_sust_df <- left_join(rsga_rgns, ao_sust_df, by="rgn_id") %>%
  group_by(iso3c) %>%
  filter(year==max(year)) %>%
  select(country = region_name, `fisheries sustainability` = score) %>%
  arrange(`fisheries sustainability`)

  datatable(ao_sust_df,
          options = list(pageLength = 10, autoWidth = TRUE),
          filter = 'top')
```

:::

## Implications

Based on the current model, this goal would be most improved by reducing the economic need for artisanal fishing in Somalia, Sudan, Dijibouti, Eritrea, Jordan, and Egypt. Access to artisanal fishing could also be improved in Eritrea.

Otherwise, access to artisanal fishing is good and artisanally fished stocks are relatively healthy. 

Future assessments should carefully assess the underlying data for improvements, here are some factors to consider:

- To determine whether artisanally fished stocks are well managed we use a global database of stock status scores that requires much gapfilling of missing data. However, regional assessments of stock status may be available.
- Do the opportunity scores seem reasonable? Should other factors be incorporated, such as piracy?
- There are better estimates of poverty than the rescaled per capita GDP used for the global assessment. For example, there may be within country estimates of the portion of population living below the poverty line.

[Other OHI+ analyses](https://oceanhealthindex.org/guidance/artisanal-fishing-opportunities/) have revised this goal to better reflect local concerns.  