---
title: ""
format: html
---

# ![](figs/biodiversity.svg){width="8%" height="8%" color="red"} Biodiversity

** This goal describes how successfully the richness and variety of marine life is being maintained. **

## Overview

People value the existence of a diverse array of species for their intrinsic qualities and their contribution to the structure and function of resilient ecosystems. The risk of species extinction generates great emotional and moral concern for many people. This concern is demonstrated by the large sums of money that people donate to help conserve species and habitats.

....
Egypt has the lowest score of 73, and the other countries score 80 or higher. These are reasonably good scores.

## Scores

```{r}
#| echo: false
#| warning: false
#| message: false

source("../_data/functions/start_script.R")
source("../_data/functions/score_display_functions.R")
score <- score(x="BD", ohi_scores)

```

::: columns
::: {.column width="50%"}
<br>

**The Biodiversity score for the Red Sea and Gulf of Aden region was `r score`.**

<span style="color: gray;">Data from the global 2023 Ocean Health Index analysis</span>
:::

::: {.column width="50%"}
```{r, fig.width=10, fig.height=3}
#| echo: false
#| warning: false
#| message: false

dial(score,colors)

```
:::
:::

The interactive map below shows the scores for the different countries.

```{r, fig.width=3, fig.height=1.5}
#| echo: false
#| warning: false
#| message: false
map_figure(ohi_scores=ohi_scores, goal="BD")

```

<br>

For the Global OHI, this goal assesses the conservation status of species based on two subgoals: Species and habitat condition. Species are assessed because they are what one typically thinks of in relation to biodiversity. We also assess Habitats because only a small proportion of marine species worldwide have been mapped and assessed, so we consider habitats a proxy for the condition of the broad suite of species that depend on them.

We calculate each of these subgoals separately and average them to estimate the overall goal score. A score of 100 means all species are at very low risk of extinction and all habitats are conserved.

The subgoal scores can be viewed below.

::: {.callout-note collapse=true}
## Subgoal scores (click to expand)

### Species condition

```{r}
#| echo: false
#| warning: false
#| message: false

source("../_data/functions/start_script.R")
#source("_data/functions/start_script.R")
source("../_data/functions/score_display_functions.R")
#source("_data/functions/score_display_functions.R")


score_spp <- score(x="SPP", ohi_scores)

```

::: columns
::: {.column width="50%"}
<br>

**The Species subgoal score for the Red Sea and Gulf of Aden region was `r score_spp`.**

<span style="color: gray;">Data from the global 2023 Ocean Health Index analysis</span>
:::

::: {.column width="50%"}
```{r, fig.width=10, fig.height=3}
#| echo: false
#| warning: false
#| message: false
source("../_data/functions/start_script.R")
#source("_data/functions/start_script.R")
source("../_data/functions/score_display_functions.R")
#source("_data/functions/score_display_functions.R")

dial(score_spp,colors)

```
:::
:::

The interactive map below shows the scores for the different countries.

```{r, fig.width=3, fig.height=1.5}
#| echo: false
#| warning: false
#| message: false

source("../_data/functions/start_script.R")
#source("_data/functions/start_script.R")
source("../_data/functions/score_display_functions.R")
#source("_data/functions/score_display_functions.R")

map_figure(ohi_scores=ohi_scores, goal="SPP")

```

<br>

### Habitat condition

```{r}
#| echo: false
#| warning: false
#| message: false

source("../_data/functions/start_script.R")
#source("_data/functions/start_script.R")
source("../_data/functions/score_display_functions.R")
#source("_data/functions/score_display_functions.R")

score_hab <- score(x="HAB", ohi_scores)

```

::: columns
::: {.column width="50%"}
<br>

**The Habitat subgoal score for the Red Sea and Gulf of Aden region was `r score_hab`.**

<span style="color: gray;">Data from the global 2023 Ocean Health Index analysis</span>
:::

::: {.column width="50%"}
```{r, fig.width=10, fig.height=3}
#| echo: false
#| warning: false
#| message: false

source("../_data/functions/start_script.R")
#source("_data/functions/start_script.R")
source("../_data/functions/score_display_functions.R")
#source("_data/functions/score_display_functions.R")

dial(score_hab,colors)

```
:::
:::

The interactive map below shows the scores for the different countries.

```{r, fig.width=3, fig.height=1.5}
#| echo: false
#| warning: false
#| message: false


source("../_data/functions/start_script.R")
#source("_data/functions/start_script.R")
source("../_data/functions/score_display_functions.R")
#source("_data/functions/score_display_functions.R")

map_figure(ohi_scores=ohi_scores, goal="HAB")

```
:::

## Model description

### Species
The species component of the biodiversity goal measures the conservation status of marine species.

This subgoal assesses the health of all marine species present in a country’s EEZ, including endangered species and species in relatively good conditions. The presence of species that are not at risk leads to a higher score.

::: {.callout-note collapse=true}
## Species calculation details (click to expand)

Species status was calculated as the area weighted average of the conservation status of the assessed species within each region. Marine species distribution and threat category data mostly came from IUCN Red List of Threatened Species, and we limited data to all species having IUCN habitat system of "marine" [http://www.iucnredlist.org](http://www.iucnredlist.org) [@iucn2022; @iucn_spatial_2022]. Seabird distributions data came from Birdlife International [http://datazone.birdlife.org](http://datazone.birdlife.org) [@birdlifeinternationalandhandbookofthebirdsoftheworld2020].   

We scaled the lower end of the biodiversity goal to be 0 when 75% species are extinct, a level comparable to the five documented mass extinctions [@barnosky_has_2011] and would constitute a catastrophic loss of biodiversity. 

Threat weights, $w_{i}$, were assigned based on the IUCN threat categories status of each $i$ species, following the weighting schemes developed by Butchart et al. [-@butchart2007improvements]. For the purposes of this analysis, we included only data for extant species for which sufficient data were available to conduct an assessment. We did not include the Data Deficient classification as assessed species following previously published guidelines for a mid-point approach [@schipper2008status; @hoffmann_impact_2010].

We first calculated each the region’s area-weighted average species risk status, $\bar R_{spp}$. For each 0.5 degree grid cell within a region, $c$, the risk status, $w$, for each species, $i$, present is summed and multiplied by cell area $A_c$, to get an area- and count-weighted species risk for each cell.  This value is then divided by the sum of count-weighted area $A_c \times N_c$ across all cells within the region.  The result is the area-weighted mean species risk across the entire region.

$$
\bar R_{spp} = \frac { \displaystyle\sum_{ c=1 }^{ M } \left( \displaystyle\sum _{ i=1 }^{N_c} w_i \right) \times A_c } 
                { \displaystyle\sum_{ c=1 }^{ M } A_c \times N_c },  (Eq. 6.4)
$$
To convert $\bar R_{spp}$ into a score, we set a floor at 25% (representing a catastrophic loss of biodiversity, as noted above) and then rescaled to produce a $x_{spp}$ value between zero and one.

$$
x_{spp} = max \left( \frac { \bar R_{SPP} - .25 }{ .75 }, 0 \right), (Eq. 6.5)
$$


**Weights for assessment of species status based on IUCN risk categories**

| Risk Category   | IUCN code | Weight|
|-----------------|-----------|-------|
| Extinct         |     EX    |  0.0  |
| Critically Endangered | CR  |  0.2  |
| Endangered      |     EN    |  0.4  |
| Vulnerable      |     VU    |  0.6  |
| Near Threatened |     NT    |  0.8  |
| Least Concern   |     LC    |  1.0  |
:::

### Habitat
The habitat component of the biodiversity goal measures the status of marine habitats that support large numbers of marine species.

Healthier habitats mean healthier species! Habitat is included as a subgoal of the Biodiversity goal to provide a more complete picture of diversity in the system than only considering the conservation status of species (the other subgoal for Biodiversity). This is because species assessments are often incomplete.

This goal assesses the condition of marine habitats that are particularly important in supporting large numbers of marine species. The status of each habitat is its current condition relative to its reference condition, which is often based on historical area. A score of 1 indicates that the condition equals or exceeds the reference point. The subgoal score is then calculated by combining the habitat scores for each region.

For the global OHI assessment, we included mangroves, coral reefs, kelp forests, seagrass beds, salt marshes, sea ice, or subtidal soft-bottom habitats. All habitats measured contribute equally to the score, regardless of their extent, because the presence of a diverse set of habitats, as well as the level of conservation of each, is considered valuable to achieve this goal.


::: {.callout-note collapse=true}
## More about global habitats (click to expand)

Habitat | Condition | Extent | Trend
-------- | --------------------- | ----------------- | ------------------
Seagrass | Increasing or stable trend assigned condition = 1.0; decreasing trend assigned condition = 0.71 based on global loss | Seagrass extent per oceanic region (vector based) | Calculated across data from 1990 - 2000
Kelp | Increasing or stable trend assigned condition = 1.0; decreasing trend assigned condition based on a 2% global yearly loss | Kelp extent per oceanic region (vector based) | Calculated across data from 1983 - 2012
Coral reefs | Current % cover divided by reference % cover | Coral reef extent per oceanic region (Vector based) | Calculated across data from 1975-2006
Mangroves | Current hectares divided by reference hectares, for coastal mangroves only | Mangrove extent per oceanic region (vector based) | Calculated using 5 most recent years of data
Salt marsh | All regions assigned condition = 0.75 based on conservative historical extent losses | Salt marsh extent per oceanic region | All regions assigned trend based on historical .28% global yearly loss 
Sea ice edge | Current (3-year rolling-average using the current year and previous 2 years) % cover of sea ice having 10-50% cover, divided by reference % cover average from 1979-2000 | Same as condition | Calculated from the fitted slope of %-deviation-from-reference per year, of the most recent 5 years of data (each year of data is based on 3-year average)
Soft bottom | Soft-bottom destructive fishing practices relative to area of soft-bottom habitat and rescaled to 0-1 based on relative global values | Halpern et al. [-@halpern2008global] | Calculated using 5 most recent years of condition data
Tidal flat | Average tidal flat extent of 2010 and 2013 relative to historic extent (average of 1989 and 1992) | Tidal flat extent per oceanic region (vector based) | Calculated across data from 2001-2013
::::

## A closer look at the data

The final score for each goal is based on its current status and future predicted status. Future status is the current status adjusted by recent trends as well as the pressures and resilience measures acting on the goal.

A great deal of data are used to calculate the final score for each goal, but the biggest factor driving scores, by far, is the current status component. Here, we take a closer look at the data underlying the status scores for each country with eez territory in the Red Sea and Gulf of Aden.

### Current status

We will first look at the data underlying the species condition subgoal and then the habitat subgoal. 

::: {.callout-note collapse=true}

## Species condition: status data (click to expand)
Getting the underlying data for this goal requires processing a large amount of spatial data.

Given this, this section is on hold for now.

```{r}
#| echo: false
#| warning: false
#| message: false

## final scores, not very helpful:
#spp <- filter(datafiles, layer=="spp_status") 

#spp_df  <- read_csv(file.path("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2023/gh-pages/",
#gsub("ohiprep:", "", spp$dir), spp$fn)) 

#############
#rgn <- terra::rast("/home/shares/ohi/git-annex/globalprep/spp/v2022/spp_risk_dists/spatial/cell_id_500m.tif")
#ocean <- terra::rast("/home/shares/ohi/git-annex/globalprep/spp/v2022/spp_risk_dists/spatial/ocean_500m.tif")
#spp <- read_csv(list.files("/home/shares/ohi/git-annex/globalprep/spp/v2022/spp_risk_dists/spp_rasters", full=TRUE)[1])  
#  tmp <- read_csv(comp_files[1])
  
#iucn <- read_csv("/home/shares/ohi/git-annex/globalprep/spp/v2022/spp_risk_dists/iucn/spp_systems_from_api_2022-1.csv") %>%
#  filter(marine_system == TRUE)

#iucn <- read_csv("/home/shares/ohi/git-annex/globalprep/spp/v2022/spp_risk_dists/iucn/spp_info_from_api_2022-1.csv") %>%
#  filter(iucn_sid == 22698175)
  
```

:::


<br>

::: {.callout-note collapse=true}

## Habitat condition: data (click to expand)
Extent reflects the estimated area of the habitat; health is the condition of the habitat on a scale from 0 to 1, with 1 indicating the current condition is at, or better, than the reference point; trend is the average yearly change in health during the most recent 5 years.  

```{r}
#| echo: false
#| warning: false
#| message: false

goal <- "hab"
rsga_rgns <- read_csv(here("_data/output/rsga_ohi_rgns.csv"))

# list of OHI datafiles:
datafiles <- read_csv("https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/metadata_documentation/layers_eez_base.csv") %>%
  filter(str_detect(layer, goal)) %>%
  filter(str_detect(layer, "hab_")) %>%
  filter(layer != "element_wts_hab_pres_abs") %>%
  filter(layer != "hab_rockyreef_extent")

all_habs <- data.frame()
for(i in 1:24){
#i=1
hab <- datafiles[i,] 
hab_df  <- read_csv(file.path("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2023/gh-pages/",
gsub("ohiprep:", "", hab$dir), hab$fn)) %>%
  select(rgn_id, year, habitat, everything()) %>%
  mutate(layer = hab$layer) 

hab_df <- hab_df %>%
  mutate(units = names(hab_df)[4])

names(hab_df)[4] <- "score"
all_habs <- rbind(all_habs, hab_df)
all_habs
  }

hab_df <- left_join(rsga_rgns, all_habs, by="rgn_id") %>%
  group_by(iso3c, habitat, units) %>%
  filter(year==max(year)) %>%
  ungroup()

hab_df1 <- hab_df %>%
  select(iso3c, country=region_name, habitat, layer, score) %>%
  mutate(score = round(score, 1)) %>%
  mutate(layer = gsub("tidal_flat", "tidal-flat", layer)) %>%
  separate(layer, into = c("first", "second", "third"), sep = "_") %>%
  select(-first, -second) %>%
  pivot_wider(names_from = third, values_from = score) %>%
  mutate(extent = ifelse(is.na(extent), 0, extent)) %>%
  mutate(health = ifelse(extent == 0 , NA, health)) %>%
  mutate(trend = ifelse(extent == 0, NA, trend)) %>%
  rename(extent_km2 = extent)


  datatable(hab_df1,
          options = list(pageLength = 10, autoWidth = TRUE),
          filter = 'top')


```
:::

## Implications

Based on the current model, this goal would be most improved by:

- reducing reliance on fishing practices that are destructive to marine soft-bottom habitats, such as dredging, particularly in Saudi Arabia
- restoring mangrove habitat

Future assessments should carefully assess the underlying data for improvements because typically better habitat data is available within each country, this allows for improved estimates of extent, condition, and trend.

The inclusion of both species and habitat levels in the model should also be discussed.

If a habitat subgoal is included, evaluate how to best weight the habitats (currently equal weighting) and whether these habitats fully capture the important habitats of the region. 

Other OHI+ analyses have revised the [habitat](https://oceanhealthindex.org/guidance/habitats/) and [species](https://oceanhealthindex.org/guidance/species/) subgoals to better reflect local concerns.  