---
title: ""
format: html
---

# ![](figs/food-provision.svg){width="8%" height="8%" color="red"} Food provision

**This goal assesses our ability to sustainably maximize wild-caught or farmed marine foods.**

## Overview

One of the most fundamental services the ocean provides is the provision of food. Fish provide nearly 20% of global animal protein, helping to meet the basic nutritional needs of over half of the world’s population (FAO 2020)!


## Scores

```{r}
#| echo: false
#| warning: false
#| message: false

source("../_data/functions/start_script.R")
source("../_data/functions/score_display_functions.R")
score <- score(x="FP", ohi_scores)

```

::: columns
::: {.column width="50%"}
<br>

**The Food Provision score for the Red Sea and Gulf of Aden region was `r score`.**

<span style="color: gray;">Data from the global 2023 Ocean Health Index analysis</span>
:::

::: {.column width="50%"}
```{r, fig.width=10, fig.height=3}
#| echo: false
#| warning: false
#| message: false

dial(score,colors)

```
:::
:::

The interactive map below shows the scores for the different countries.

```{r, fig.width=3, fig.height=1.5}
#| echo: false
#| warning: false
#| message: false
map_figure(ohi_scores=ohi_scores, goal="FP")

```

<br>

For the Global OHI, this goal assesses seafood obtained from wild-caught fisheries and marine farming, or mariculture. 

We calculate each of these subgoals separately and average them to estimate the overall goal score. For this goal we weight the contributions to the final score based on the tonnes of production. A score of 100 means that food production is *sustainably* maximized.

Mariculture scores are based on two factors:

- Maximizing the amount of seafood that can grown based on conditions.
- Ensuring that mariculture practices are sustainable.

The subgoal scores can be viewed below.

:::{.callout-note collapse=TRUE}


## Subgoal scores (click to expand)

### Mariculture

```{r}
#| echo: false
#| warning: false
#| message: false

source("../_data/functions/start_script.R")
#source("_data/functions/start_script.R")
source("../_data/functions/score_display_functions.R")
#source("_data/functions/score_display_functions.R")


score_mar <- score(x="MAR", ohi_scores)

```

::: columns
::: {.column width="50%"}
<br>

**The Mariculture subgoal score for the Red Sea and Gulf of Aden region was `r score_mar`.**

<span style="color: gray;">Data from the global 2023 Ocean Health Index analysis</span>
:::

::: {.column width="50%"}
```{r, fig.width=10, fig.height=3}
#| echo: false
#| warning: false
#| message: false
source("../_data/functions/start_script.R")
#source("_data/functions/start_script.R")
source("../_data/functions/score_display_functions.R")
#source("_data/functions/score_display_functions.R")

dial(score_mar,colors)

```
:::
:::

The interactive map below shows the scores for the different countries.

```{r, fig.width=3, fig.height=1.5}
#| echo: false
#| warning: false
#| message: false

source("../_data/functions/start_script.R")
#source("_data/functions/start_script.R")
source("../_data/functions/score_display_functions.R")
#source("_data/functions/score_display_functions.R")

map_figure(ohi_scores=ohi_scores, goal="MAR")

```

<br>

### Wildcaught fisheries

```{r}
#| echo: false
#| warning: false
#| message: false

source("../_data/functions/start_script.R")
#source("_data/functions/start_script.R")
source("../_data/functions/score_display_functions.R")
#source("_data/functions/score_display_functions.R")

score_hab <- score(x="FIS", ohi_scores)

```

::: columns
::: {.column width="50%"}
<br>

**The Wildcaught Fisheries subgoal score for the Red Sea and Gulf of Aden region was `r score_hab`.**

<span style="color: gray;">Data from the global 2023 Ocean Health Index analysis</span>
:::

::: {.column width="50%"}
```{r, fig.width=10, fig.height=3}
#| echo: false
#| warning: false
#| message: false

source("../_data/functions/start_script.R")
#source("_data/functions/start_script.R")
source("../_data/functions/score_display_functions.R")
#source("_data/functions/score_display_functions.R")

dial(score_hab,colors)

```
:::
:::

The interactive map below shows the scores for the different countries.

```{r, fig.width=3, fig.height=1.5}
#| echo: false
#| warning: false
#| message: false


source("../_data/functions/start_script.R")
#source("_data/functions/start_script.R")
source("../_data/functions/score_display_functions.R")
#source("_data/functions/score_display_functions.R")

map_figure(ohi_scores=ohi_scores, goal="FIS")

```
:::


## Model description
The Food Provision goal measures the amount of seafood sustainably harvested through any means for human consumption, and thus includes both mariculture and wild-caught fisheries (commercial, artisanal, and recreational).For the global assessment, the contribution of these two subgoals to the overall score is weighted by their total seafood yield. However, other approaches for weighting the contribution of these subgoals are reasonable.

A high score indicates that a region has maximized the amount of sustainably produced seafood from wild or ocean-cultured stocks. For fisheries, both over- and under- harvesting should be penalized. For mariculture, production should be at the sustainable “potential”. 


### Mariculture

As humanity’s population continues to expand, we must increasingly rely upon new and improved methods for cultivating food on land and in the ocean. The term aquaculture refers to the cultivation in water (fresh and saltwater) of animals or plants for human consumption. Mariculture refers to the commercial harvest of seafood that is farm-raised in the ocean and along the coast.

Aquaculture is the fastest growing animal food production sector in the world, with a 527% rise in production from 1990 to 2018 (FAO 2020)! As the population contines to grow, it will become increasingly important to efficiently grow food using sustainable mariculture practices.

::: {.callout-note collapse="true"}
## More details (click to expand)
The status of the mariculture subgoal, $x_{mar}$, was calculated using three sources of data:

- *Mariculture production:* Tonnes of production for both marine and brackish waters, excluding species that were not used as a source of food for human consumption is from FAO.

- *Potential production:* Gentry et al. (2019) estimates the tonnes of possible production for marine aquaculture at a high resolution spatial scale. To account for social and economic realities and other constraints to these ecological potentials, we constrained the per-country potential to 1% of this tonnage estimate and used these country values as reference points. 

- *Sustainability:* Additionally, we include a sustainability score for each species in each region which is based on the [Monterey Bay Aquarium Seafood Watch aquaculture recommendations](https://www.seafoodwatch.org/seafood-recommendations/standards-revision) [@2023].

:::

### Wild caught fisheries

The wild-caught fisheries subgoal evaluates how well we maximize harvest without damaging the ocean’s ability to continue providing fish for people in the future. The sustainable harvest of wild-caught seafood avoids excessively high exploitation of target species and does not target threatened populations. Each year’s catch must not take so much from the population that future catches will decline. 

::: {.callout-note collapse="true"}
## More details (click to expand)
We assess food provision from wild caught fisheries by estimating population biomass relative to the biomass that can deliver maximum sustainable yield ($B/B_{MSY}$) for each stock. When available, we obtained $B/B_{MSY}$ values from the RAM Legacy Stock Assessment Database.

When RAM data were not available, we used data-limited approaches that have been developed to estimate $B/B_{MSY}$ values using globally available catch data. 

To calculate the status for each region and year, $B/B_{MSY}$ values were converted to a stock status score between 0-1 that penalizes over-harvesting. To obtain the overall status for each region, the stock status scores for all the stocks within a region were averaged using a geometric mean weighted by the average catch (tonnes) of each stock using Sea Around Us catch data. 

:::


## A closer look at the data

The biggest factor driving scores, by far, is the current status component. Here, we take a closer look at the data underlying the status scores for each country with EEZ territory in the Red Sea and Gulf of Aden.

### Current status

We will first look at the data underlying the mariculture subgoal and then then wildcaught fisheries subgoal. 


::: {.callout-note collapse=true}

## Mariculture: status data (click to expand)
The tonnes of production and sustainability of rearing practices for each species that is currently grown in each country.Sustainability is rescaled to values between 0 and 1, with 1 being perfectly sustainable.



```{r}
#| echo: false
#| warning: false
#| message: false

goal <- "mar_"
rsga_rgns <- read_csv(here("_data/output/rsga_ohi_rgns.csv"))

# list of OHI datafiles:
datafiles <- read_csv("https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/metadata_documentation/layers_eez_base.csv") %>%
  filter(str_detect(layer, goal)) 

# prep files:
prep <- read_csv("https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/metadata_documentation/layers_eez_dataprep.csv")

```


```{r}
#| echo: false
#| warning: false
#| message: false

# "mar_harvest_tonnes", "mar_sustainability_score", "mar_capacity"  
tonnes <- filter(datafiles, layer=="mar_harvest_tonnes") 

tonnes_df  <- read_csv(file.path("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2023/gh-pages/", gsub("ohiprep:", "", tonnes$dir), tonnes$fn)) 

tonnes_df <- left_join(rsga_rgns, tonnes_df, by="rgn_id") %>%
   filter(year==2021) %>%
  filter(tonnes > 0)
  

systainability_file <- filter(datafiles, layer=="mar_sustainability_score") 

sustainability_df  <- read_csv(file.path("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2023/gh-pages/", gsub("ohiprep:", "", systainability_file$dir), systainability_file$fn)) %>%
  select(rgn_id, taxa_code, sustainability=sust_coeff)

mar_df <- tonnes_df %>%
  left_join(sustainability_df, by=c("rgn_id", "taxa_code") ) %>%  mutate(taxa_code = str_replace(taxa_code, "_\\d+", "")) %>%
  select(iso3c, country=region_name, species=taxa_code, tonnes, sustainability)

  datatable(mar_df,
          options = list(pageLength = 10, autoWidth = TRUE),
          filter = 'top')


```


The following table describes the tonnes of production relative to the tonnes of production that is possible.
```{r}
#| echo: false
#| warning: false
#| message: false

#tonnes relative to capacity
mar_total <- tonnes_df %>%
  group_by(iso3c) %>%
  summarize(production_tonnes = sum(tonnes))

potential_file <- filter(datafiles, layer=="mar_capacity") 

potential_df  <- read_csv(file.path("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2023/gh-pages/", gsub("ohiprep:", "", potential_file$dir), potential_file$fn)) 

potential_df <- left_join(rsga_rgns, potential_df, by="rgn_id") %>%
  select(iso3c, country=region_name, potential_tonnes=potential_mar_tonnes) %>%
  left_join(mar_total, by="iso3c") %>%
  mutate(production_tonnes = ifelse(is.na(production_tonnes), 0, production_tonnes)) %>%
  mutate(proportion = round(production_tonnes/potential_tonnes, 3)) %>%
  mutate(proportion = ifelse(production_tonnes == 0 & potential_tonnes==0, 0, proportion)) %>%
  select(iso3c, country, production=production_tonnes, potential=potential_tonnes, proportion)

  datatable(potential_df,
          options = list(pageLength = 10, autoWidth = TRUE),
          filter = 'top')
```

:::


<br>


::: {.callout-note collapse=true}

## Wildcaught fisheries: data (click to expand)

This table describes recorded fisheries catch with corresponding stock status scores (i.e., bbmsy). B/Bmsy scores of 1, or greater, indicate the stock is healthy and current biomass is at, or above, maximum sustainable yield.

```{r}
#| echo: false
#| warning: false
#| message: false

goal <- "fis"
rsga_rgns <- read_csv(here("_data/output/rsga_ohi_rgns.csv"))

# list of OHI datafiles:
datafiles <- read_csv("https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/metadata_documentation/layers_eez_base.csv") %>%
  filter(str_detect(layer, goal))

#status
status_file <- filter(datafiles, layer=="fis_b_bmsy") 

status_df  <- read_csv(file.path("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2023/gh-pages/", gsub("ohiprep:", "", status_file$dir), status_file$fn)) 

status_df <- left_join(rsga_rgns, status_df, by="rgn_id") %>%
   filter(year==2019) %>%
  select(iso3c, stock_id,  bbmsy)


# catch
catch_file <- filter(datafiles, layer=="fis_meancatch") 

catch_df  <- read_csv(file.path("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2023/gh-pages/", gsub("ohiprep:", "", catch_file$dir), catch_file$fn)) 

catch_df <- left_join(rsga_rgns, catch_df, by="rgn_id") %>%
   filter(year==2019) %>%
  mutate(stock_id = str_replace(stock_id_taxonkey, "_[0-9]+$", "")) %>%
  select(iso3c, country=region_name, stock_id, mean_catch)

# score data
fish_df <- left_join(catch_df, status_df, by=c("iso3c", "stock_id")) %>%
  group_by(iso3c) %>%
  arrange(iso3c, -mean_catch)
# 
# fish_df %>%
#   group_by(iso3c) %>%
#   summarize(mean_bbmsy = mean(bbmsy, na.rm=TRUE))
#                           
#   fish_df %>%
#   mutate(status_score = ifelse(is.na(bbmsy), 0, 1)) %>%
#   group_by(status_score) %>%
#   summarize(catch = sum(mean_catch))
                      
  datatable(fish_df,
          options = list(pageLength = 10, autoWidth = TRUE),
          filter = 'top')

```
:::


## Implications

Overall, both mariculture and fisheries scores are low for all the countries in this region.

Based on the current model, mariculture scores are very low primarily because production is low compared to the potential for production. About 226,000 tonnes of seafood are farmed each year in this region, with Egypt producing the most (~138,00 tonnes) and Saudi Arabia next (~85,000). However, this is small relative to the 3.7 million tonnes that is could, theoretically, be produced.

The sustainability scores for reared seafood are fairly high for all countries and taxa, with the exception of Egypt. Egypt has low sustainability for European seabass, Gilthead seabream, and Meagre.

Fisheries scores are low primarily because most of the wild-caught fish from this region do not have stock stack scores, which we use as an indicator of weak fisheries management for the global assessment. Only about 40% of the fish catch is identified to species, which means that it can not be assessed using stock assessments and this is heavily penalized in the global Ocean Health Index fisheries model. 

Future assessments should carefully assess the underlying data and models to meet their objectives.

Other OHI+ analyses have revised [mariculture](https://oceanhealthindex.org/guidance/mariculture/) and [fisheries](https://oceanhealthindex.org/guidance/wild-caught-fisheries/) subgoals to better reflect local concerns.  
