---
title: ""
format: html
---
# ![](figs/natural-products.svg){width=8% height=8% color="red"} Natural products 

## Overview
This goal assesses how well countries are maximizing the sustainable harvest of non-food marine resources.

Marine natural products can provide substantial economic benefits when sustainably harvested and traded. For example, millions of marine ornamental (aquarium) fish are harvested and exported for sale each year. This trade can provide a valuable and long-term income stream for low-income coastal communities if undertaken in a sustainable manner.

## Scores
```{r}
#| echo: false
#| warning: false
#| message: false

source("../_data/functions/start_script.R")
source("../_data/functions/score_display_functions.R")
score <- score(x="NP", ohi_scores)

```

:::: {.columns} 

::: {.column width="50%"}

<br>
<br>
Using data from the global 2023 Ocean Health Index analysis, the overall score for the Red Sea and Gulf of Aden region was `r score`.
:::

::: {.column width="50%"}

```{r, fig.width=10, fig.height=3}
#| echo: false
#| warning: false
#| message: false

dial(score,colors)

```

:::

::::

The interactive map below shows how different countries contributed to this score.

```{r, fig.width=3, fig.height=1.5}
#| echo: false
#| warning: false
#| message: false
map_figure(ohi_scores=ohi_scores, goal="NP")

```

<br>


## Model description
This goal score is based on the sustainable harvest quantity, relative to some target, of marine natural products such as seaweed and marine plants, sponges, shells, fish oil and meal, and marine ornamental fish. This goal does not include bioprospecting which focuses on potential (and largely unknowable and potentially infinite) value rather than current realized value, or non-living products such as oil and gas or mining products which by definition are not sustainable. Ideally targets are based on the amount of harvest that does not threaten future harvest, such as the maximum sustainable yield from fisheries stock status scores.

Currently, the global Ocean Health Index includes harvest and production of non-food seaweed, ornamental fish, and wild-caught fish used for fish oil and meal.

A high score indicates that natural resources are maximally harvested at a rate that does not threaten future harvest and minimizes ecological harm.

## Regional analysis

More details to come!
```{r}

```