---
title: ""
format: html
---

# ![](figs/sense-of-place.svg){width="8%" height="8%" color="red"} Sense of place

**This goal describes how well we are preserving current and future access to coastal and marine systems that people value as part of their cultural identity.**

## Overview

The ocean is an important source of our collective cultural identity. Our connection to marine and coastal environments is profound, providing inspiration and a sense of history and place. For the global assessment, we measure how well we are protecting this legacy for future generations by assessing the condition of iconic species (Iconic Species subgoal) and the protection of coastal regions (Lasting Special Places subgoal).


## Scores

```{r}
#| echo: false
#| warning: false
#| message: false

source("../_data/functions/start_script.R")
source("../_data/functions/score_display_functions.R")
score <- score(x="SP", ohi_scores)

```

::: columns
::: {.column width="50%"}
<br>

**The Sense of Place score for the Red Sea and Gulf of Aden region was `r score`.**

<span style="color: gray;">Overall, the scores for this goal are low throughout the region. This suggests that the marine places people value may not be well protected for future generations. </span>

:::

::: {.column width="50%"}
```{r, fig.width=10, fig.height=3}
#| echo: false
#| warning: false
#| message: false

dial(score,colors)

```
:::
:::

The interactive map below shows the scores for the different countries.

```{r, fig.width=3, fig.height=1.5}
#| echo: false
#| warning: false
#| message: false
map_figure(ohi_scores=ohi_scores, goal="SP")

```

<br>

For the Global OHI, this goal captures the aspects of the coastal and marine system that people value as part of their cultural identity by measuring two subgoals: the condition of iconic species and the condition of lasting special places.

We calculate each of these subgoals separately and average them to estimate the overall goal score. A score of 100 means all iconic species are at very low risk of extinction and at least 30% of coastal habitats are designated as marine protected areas.

The subgoal scores can be viewed below.

::: {.callout-note collapse=true}
## Subgoal scores (click to expand)

### Iconic species condition

```{r}
#| echo: false
#| warning: false
#| message: false

source("../_data/functions/start_script.R")
#source("_data/functions/start_script.R")
source("../_data/functions/score_display_functions.R")
#source("_data/functions/score_display_functions.R")


score_spp <- score(x="ICO", ohi_scores)

```

::: columns
::: {.column width="50%"}
<br>

**The Iconic species subgoal score for the Red Sea and Gulf of Aden region was `r score_spp`.**

<span style="color: gray;">This goal is low for all countries which indicates that too many of the marine animals that people value are critically endangered or endangered.</span>

:::

::: {.column width="50%"}
```{r, fig.width=10, fig.height=3}
#| echo: false
#| warning: false
#| message: false
source("../_data/functions/start_script.R")
#source("_data/functions/start_script.R")
source("../_data/functions/score_display_functions.R")
#source("_data/functions/score_display_functions.R")

dial(score_spp,colors)

```
:::
:::

The interactive map below shows the scores for the different countries.

```{r, fig.width=3, fig.height=1.5}
#| echo: false
#| warning: false
#| message: false

source("../_data/functions/start_script.R")
#source("_data/functions/start_script.R")
source("../_data/functions/score_display_functions.R")
#source("_data/functions/score_display_functions.R")

map_figure(ohi_scores=ohi_scores, goal="ICO")

```


<br>

### Lasting special places

```{r}
#| echo: false
#| warning: false
#| message: false

source("../_data/functions/start_script.R")
#source("_data/functions/start_script.R")
source("../_data/functions/score_display_functions.R")
#source("_data/functions/score_display_functions.R")

score_hab <- score(x="LSP", ohi_scores)

```

::: columns
::: {.column width="50%"}
<br>

**The overall Lasting Special Places subgoal score for the Red Sea and Gulf of Aden region was `r score_hab`.**

<span style="color: gray;">The low scores for this goal suggest little effort has been made to preserve culturally valuable coastal and marine areas.</span>

:::

::: {.column width="50%"}
```{r, fig.width=10, fig.height=3}
#| echo: false
#| warning: false
#| message: false

source("../_data/functions/start_script.R")
#source("_data/functions/start_script.R")
source("../_data/functions/score_display_functions.R")
#source("_data/functions/score_display_functions.R")

dial(score_hab,colors)

```
:::
:::

The interactive map below shows the scores for the different countries.

```{r, fig.width=3, fig.height=1.5}
#| echo: false
#| warning: false
#| message: false


source("../_data/functions/start_script.R")
#source("_data/functions/start_script.R")
source("../_data/functions/score_display_functions.R")
#source("_data/functions/score_display_functions.R")

map_figure(ohi_scores=ohi_scores, goal="LSP")

```
:::


## Model description

### Iconic species
The Iconic species subgoal measures the  status of iconic marine species, which are the animals that have unique importance to humans as demonstrated through traditional activities, ethnic or religious practices, existence value, or locally acknowledged aesthetic value.

Iconic species are those that are relevant to local cultural identity through a species’ relationship to one or more of the following: 1) traditional activities such as fishing, hunting or commerce; 2) local ethnic or religious practices; 3) existence value; and 4) locally-recognized aesthetic value (e.g., touristic attractions/common subjects for art such as whales).

This sub-goal is estimated by averaging the status scores of the iconic species in each region based on their IUCN Red List threat categories. 

The list of iconic species was drawn from several data sources, but primarily from the World Wildlife Fund’s global and regional lists for Priority Species (especially important to people for their health, livelihoods, and/or culture) and Flagship Species (‘charismatic’ and/or well-known). Additional culturally important species species, available at the continent level (Garcia et al. 2023), were added to supplement the original iconic species list. 


### Lasting special places
The Lasting special places sub-goal measures how well we are protecting, for current and future generations, the locations that contribute to marine-related cultural identity.

Special places are coastal and marine areas that are culturally significant for a variety of reasons. Designating a place to be a landmark, monument, and/or Marine Protected Area can help ensure its existence, provide valuable ecosystem services and preserve cultural heritage and identity.

To get a score of 100 a region must protect at least 30% of its coastal area. 

We use the United Nation’s World Database on Protected Areas (WDPA) to identify protected areas. The WDPA aggregates several key databases: IUCN’s World Commission on Protected Areas, Global Marine Protected Areas, UNESCO World Heritage Marine sites, National Parks and Nature Reserves, and the United Nations List of Protected Places.  In most cases the year of designation is listed for each protected area.

We focus only on coastal waters (within 3 nautical miles of shore) for marine special places because we assume lasting special places are primarily in coastal areas. For coastlines, we focus only on the first 1-km-wide strip of land as a way to increase the likelihood that the area being protected by terrestrial parks is connected to the marine system in some way.


::: {.callout-note collapse=true}
## More details (click to expand)

The lasting special places sub-goal focuses on geographic locations that hold particular value for aesthetic, spiritual, cultural, recreational or existence reasons. This sub-goal is particularly hard to quantify. Ideally one would survey every community around the world to determine the top list of special places, and then assess how those locations are faring relative to a desired state (e.g., protected or well managed). The reality is that such lists do not exist. Instead, we assume areas that are protected indicate special places (i.e., the effort to protect them suggests they are important places). Clearly this is an imperfect assumption but in many cases it will be true.
:::

## A closer look at the data

The biggest factor driving scores, by far, is the current status component. Here, we take a closer look at the data underlying the status scores for each country with EEZ territory in the Red Sea and Gulf of Aden.

### Current status

We will first look at the data underlying the Iconic species subgoal and then the Lasting special places subgoal. 

#### Iconic species
The following is a list of iconic species found in the Red Sea and Gulf of Aden region as well as their IUCN conservation status. Iconic species include animals such as the Blue Whale (endangered), Grey Reef Shark (endangered), Blacktip Reef Shark (vulnerable), and the Humphead Wrasse (endangered). 

We identified 24 iconic species within the region. For each species we determine the IUCN threatened status designations within each region the species is located. In 52 (of N=141) cases, the species were critically endangered or endangered; 45 were vulnerable, and 34 were either of least concern or not threatened.


::: {.callout-note collapse=true}

## Iconic species condition scores (click to expand)

```{r}
#| echo: false
#| warning: false
#| message: false

goal <- "ico"
rsga_rgns <- read_csv(here("_data/output/rsga_ohi_rgns.csv"))

# list of OHI datafiles:
ico_files <- read_csv("https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/metadata_documentation/layers_eez_base.csv") %>%
  filter(str_detect(layer, goal)) 

ico_df  <- read_csv(file.path("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2023/gh-pages/",
gsub("ohiprep:", "", ico_files$dir), ico_files$fn)) %>%
  filter(year == 2022) 

ico <- left_join(rsga_rgns, ico_df, by="rgn_id") %>%
  select(iso3c, country = region_name, species = sciname, year = eval_yr, status = category) %>%
  mutate(status = case_when(
    status == "CR" ~ "Critically threatened",
    status == "EN" ~ "Endangered",
    status == "LC" ~ "Least concern",
    status == "NT" ~ "Not threatened",
    status == "VU" ~ "Vulnerable",
    TRUE ~ status  # Keeps the original value if none of the above conditions are met
  )) 

  datatable(ico,
          options = list(pageLength = 10, autoWidth = TRUE),
          filter = 'top')


```

:::


<br>

#### Lasting special places
These data describe the km2 and percent protected areas located in the 1km inland coastal areas and 3nm offshore regions. 

In regard to protected areas, only Saudi Arabia (~ 12% protected), Israel (~4%), and Egypt (~21%) have marine protected areas along the coastline and offshore regions (within 3nm of the coast), but none are near the 30% required to obtain a score of 100. 
::: {.callout-note collapse=true}

## Protected areas: Country data (click to expand)

```{r}
#| echo: false
#| warning: false
#| message: false

goal <- "lsp_"
rsga_rgns <- read_csv(here("_data/output/rsga_ohi_rgns.csv"))

# list of OHI datafiles:
datafiles <- read_csv("https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/metadata_documentation/layers_eez_base.csv") 

inland_mpa <- filter(datafiles, layer=="lsp_prot_area_inland1km") 
inland_mpa_df  <- read_csv(file.path("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2023/gh-pages/",
gsub("ohiprep:", "", inland_mpa$dir), inland_mpa$fn)) %>%
  filter(year==2022)

marine_mpa <- filter(datafiles, layer=="lsp_prot_area_offshore3nm") 
marine_mpa_df  <- read_csv(file.path("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2023/gh-pages/",
gsub("ohiprep:", "", marine_mpa$dir), marine_mpa$fn)) %>%
  filter(year==2022)

marine_area <- filter(datafiles, layer=="rgn_area_offshore3nm") 
marine_area_df  <- read_csv(file.path("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2023/gh-pages/",
gsub("ohiprep:", "", marine_area$dir), marine_area$fn)) %>%
  select(rgn_id, marine_area=area)

inland_area <- filter(datafiles, layer=="rgn_area_inland1km") 
inland_area_df  <- read_csv(file.path("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2023/gh-pages/",
gsub("ohiprep:", "", inland_area$dir), inland_area$fn)) %>%
  select(rgn_id, inland_area=area)

lsp <- left_join(rsga_rgns, marine_mpa_df, by="rgn_id") %>%
  left_join(marine_area_df, by="rgn_id") %>%
  mutate(marine_proportion = round(a_prot_3nm/marine_area*100, 0)) %>%
  left_join(inland_mpa_df, by="rgn_id") %>%
  left_join(inland_area_df, by="rgn_id") %>%
  mutate(coastal_proportion = round(a_prot_1km/inland_area*100, 0)) %>%
  select(iso3c, region_name, km2_protected_marine=a_prot_3nm, percent_marine=marine_proportion, km2_protected_coast = a_prot_1km, percent_coast = coastal_proportion)


  datatable(lsp,
          options = list(pageLength = 10, autoWidth = TRUE),
          filter = 'top')


```
:::

## Implications

Overall, more could be done to protect the species and areas that people value in this region.

Iconic species scores are fairly low throughout the region, indicating that the welfare of many iconic species, such as the Blue Whale and Green Sea Turtle, is threatened.

Furthermore, according to the global data, there are very few Marine Protected Areas in the region. Implementing Marine Protected Areas along the coastline and nearshore areas in all the countries would immediately improve scores.

Other OHI+ analyses have revised the [iconic species](https://oceanhealthindex.org/guidance/iconic-species/) and [lasting special places](https://oceanhealthindex.org/guidance/lasting-special-places/) subgoals to better reflect local concerns.  