---
title: ""
format: html
---
# ![](figs/tourism-and-recreation.svg){width=8% height=8% color="red"} Tourism and recreation 

## Overview
This goal captures the number of people, and the quality of their experience, visiting coastal and marine areas and attractions.

Tourism and recreation in coastal areas is a major component of thriving coastal communities and a measure of how much people value ocean systems. Tourism and recreation provide opportunities for the enjoyment of marine and coastal environments. The ability to visit culturally or aesthetically pleasing places with healthy coastal ecosystems, diverse wildlife, and opportunities for exploration and recreation contributes to the well-being of many people.


## Scores
```{r}
#| echo: false
#| warning: false
#| message: false

source("../_data/functions/start_script.R")
source("../_data/functions/score_display_functions.R")
score <- score(x="TR", ohi_scores)

```

:::: {.columns} 

::: {.column width="50%"}

<br>
<br>
Using data from the global 2023 Ocean Health Index analysis, the overall score for the Red Sea and Gulf of Aden region was `r score`.
:::

::: {.column width="50%"}

```{r, fig.width=10, fig.height=3}
#| echo: false
#| warning: false
#| message: false

dial(score,colors)

```

:::

::::

The interactive map below shows how different countries contributed to this score.

```{r, fig.width=3, fig.height=1.5}
#| echo: false
#| warning: false
#| message: false
map_figure(ohi_scores=ohi_scores, goal="TR")

```

<br>


## Model description

The Tourism & Recreation goal measures how successfully a region’s marine and coastal areas maintain their attractiveness as destinations for people to experience and enjoy priced activities, such as paid boat excursions, visits to coastal parks with entrance fees and ocean-view hotels, as well as unpriced activities such as free days at the beach and hikes on public trails.

Although coastal tourism industries can be important contributors to coastal economies, the tourism and recreation goal is assessed separately from its economic benefits, which are reported in the coastal livelihoods and economies goal. 

At the global scale we used international tourist arrivals as a reasonable proxy measure for the total number of people engaged in coastal tourism and recreation activities. We have not been able to incorporate domestic tourism, but we operate under the assumption that international arrivals capture the relative participation in tourist activities well enough. 

This goal also incorporates a measure of sustainability from the World Economic Forum's Travel & Tourism Development Index (TTDI). 

The reference point is relative to other countries, such that to get a score of 100, a country has to have rates of tourism that are better than (or equal to), 90% of countries.


::: {.callout-note collapse=true}
## More details (click to expand)

The model for the status of the tourism & recreation goal, $x_{tr}$, is:

$$
x_{tr} = \frac{ T_r }{ T_{90th}}, (Eq. 6.20)
$$

where, $T_{90th}$ is the $T_r$ value of the region value that corresponds to the 90th quantile, and:  

$$
T_{r} = { A }\times { S }, (Eq. 6.21)
$$

where, $A$ is the proportion of overnight tourist arrivals divided by the total international arrivals, and $S$ is sustainability. 

Ideally there would be data available specifically for arrivals, $A$, related to coastal tourism; however, the best data available at a global scale reports the total international arrivals, which does not solely reflect coastal tourism.

Unfortunately it was not possible to determine the proportion of international arrivals affiliated with strictly leisure tourism. However, some (unknown) proportion of business travelers also enjoy the coast for leisure during their visit to coastal areas, such that we assumed all tourist arrivals were related to tourism and recreation values. Regional applications of the Index can make use of better-resolved data and more direct measures of tourism, as has been done within the US West Coast, where data for participation in coastal recreational activities across 19 different sectors were available.

Measures of sustainability are data from the World Economic Forum's Travel & Tourism Development Index (TTDI). This index measures “the set of factors and policies that enable the sustainable and resilient development of the Travel and Tourism (T&T) sector, which in turn contributes to the development of a country.” The index is comprised of five subindexes 17 pillars and 112 individual indicators, distributed among the different pillars. We use scores for the Travel and Tourism Sustainability Subindex which encompasses three pillars:

    
Pillar 15: **Environmental Sustainability**

  * Greenhouse gas (GHG) emissions per capita
  * Renewable energy
  * Global Climate Risk Index
  * Investment in green energy and infrastructure
  * Particulate matter (2.5) concentration
  * Baseline water stress
  * Red List Index
  * Forest cover loss
  * Wastewater treatment
  * Clean ocean water
  * Number of environmental treaty ratifications
  * Adequate protection for nature
  * Oversight of production impact on the environment and nature
  * Total protected areas coverage
  * Average proportion of key bio
    
Pillar 16: **Socioeconomic Resilience and Conditions**

  * Poverty rate
  * Social protection basic coverage
  * Social protection spending
  * Not in education, employment or training (NEET) ratio
  * Equal workforce opportunities
  * Workers’ rights
  * Gender Inequality Index

Pillar 17: **Travel and Tourism Demand Pressure and Impact**

  * T&T GDP multiplier
  * Inbound length of stay
  * Seasonality of international tourist arrivals
  * Concentration of interest in cultural attractions
  * Concentration of interest in nature attractions
  * Geographically dispersed tourism
  * Quality of town and city centre
    
The sustainability factor, $S$, is the Travel and Tourism Sustainability Subindex score, which is the unweighted average of its three component pillars. 
:::


## Regional analysis

More details to come!
```{r}

```