---
title: "Spatial Prep"
output: html_document
date: "2023-11-14"
---

Download: https://www.marineregions.org/downloads.php
Marine and land zones: the union of world country boundaries and EEZ's
Version 3 - 2020-03-17 (19 MB) [Known issues] (created from EEZ version 11)
To cite this product:

Flanders Marine Institute (2020). Union of the ESRI Country shapefile and the Exclusive Economic Zones (version 3). Available online at https://www.marineregions.org/. https://doi.org/10.14284/403

EEZs only:
Flanders Marine Institute (2023). Maritime Boundaries Geodatabase: Maritime Boundaries and Exclusive Economic Zones (200NM), version 12. Available online at https://www.marineregions.org/. https://doi.org/10.14284/632

Downloaded: Nov 13 2023, MRF

```{r setup, include=FALSE}
library(sf)
library(tidyverse)
library(here)

mazu_spatial <- "/home/shares/ohi/OHI_RSGA/_spatial"
```
extreme sea level: https://www.nature.com/articles/s41598-020-67736-6

EEZ and country boundaries

The Hala'ib Triangle is a disputed area located on the Red Sea's African coast, between the border of Egypt and Sudan. This triangular piece of land, which spans an area of around 20,580 square kilometers (7,950 square miles), has been a subject of territorial dispute between Egypt and Sudan for many years.

The dispute over the Hala'ib Triangle can be traced back to the colonial era. In 1899, the Anglo-Egyptian Condominium Agreement for Sudan set the 22nd parallel as the boundary between Egypt and Sudan. However, in 1902, a separate administrative boundary was drawn, which placed the Hala'ib Triangle under Sudanese administration while keeping it within Egyptian territory. This delineation was mainly for administrative convenience, but it led to the current conflicting claims.

Egypt bases its claim on the original 1899 border, which puts the Hala'ib Triangle within its territory. On the other hand, Sudan claims the area based on the 1902 administrative boundary. The issue remained relatively dormant until the late 20th century, when potential offshore oil reserves in the Red Sea heightened the stakes of the territorial dispute.

Both nations have attempted to assert sovereignty over the area at various times, with Egypt currently exercising de facto control over the region. Despite this, the dispute remains unresolved, and it is a point of tension in Egyptian-Sudanese relations. International efforts to mediate the dispute have so far been unsuccessful, and the future status of the Hala'ib Triangle remains uncertain.

The Doumeira Islands, including Doumeira Island and several smaller islets, along with the nearby Doumeira Peninsula, are at the center of a territorial dispute between Eritrea and Djibouti. This dispute has led to military confrontations between the two countries, contributing to the tension in the Horn of Africa region.

Geographic Location: The Doumeira Islands are located in the Bab-el-Mandeb Strait, a strategic waterway connecting the Red Sea to the Gulf of Aden. This strategic location is a key factor in the dispute, as control over the islands would offer significant strategic and economic advantages.

Historical Background: The origin of the dispute dates back to the colonial era when Eritrea was an Italian colony, and Djibouti was under French control. The boundary between the two territories was never clearly demarcated, leading to ambiguity over the sovereignty of the Doumeira Islands and Peninsula.

Conflict and Tensions: Tensions escalated into open conflict on several occasions, most notably in 1996 and 2008. The 2008 conflict resulted in several days of fighting, causing casualties and displacements. These clashes highlighted the potential of the dispute to destabilize the broader region.

International Mediation and Current Status: Efforts to resolve the dispute have involved international mediation. After the 2008 conflict, Qatar intervened to mediate and both countries agreed to a Qatari military presence in the disputed area as a peacekeeping force. However, in 2017, following the Gulf diplomatic crisis, Qatar withdrew its troops, leading to renewed tensions.

Continuing Dispute: Despite these mediation efforts, the territorial dispute over the Doumeira Islands remains unresolved. The area continues to be a point of contention between Eritrea and Djibouti, with occasional military movements and rhetoric indicating ongoing tensions.

The dispute over the Doumeira Islands is a complex issue involving historical, strategic, and nationalistic elements. It is illustrative of the broader challenges facing the Horn of Africa, a region marked by border disputes, political instability, and strategic competition.

Cutting the persian gulf portion of Saudi Arabia's EEZ.



```{r}

eez_land <- st_read(file.path(mazu_spatial, "data/EEZ_land_union_v3_202003/EEZ_Land_v3_202030.shp"))
test <- filter(eez_land, UNION=="United States")
plot(st_geometry(test))

red_sea_major <- st_read(file.path(mazu_spatial, "data/red_sea/red_sea.shp"))
plot(st_geometry(red_sea), add=TRUE)
red_sea_suez <- st_read(file.path(mazu_spatial, "data/suez/suez.shp"))
plot(st_geometry(red_sea_suez), add=TRUE)
red_sea_aqaba <- st_read(file.path(mazu_spatial, "data/Aqaba/Aqaba.shp"))
plot(st_geometry(red_sea_aqaba), add=TRUE)

red_sea_tmp <- rbind(red_sea_major, red_sea_aqaba, red_sea_suez)
red_sea_tmp <- st_union(red_sea_tmp)
red_sea_tmp <- st_sf(red_sea_tmp)
red_sea_tmp <- red_sea_tmp[st_geometry_type(red_sea_tmp) == "POLYGON" | st_geometry_type(red_sea_tmp) == "MULTIPOLYGON", ]
red_sea <- red_sea_major
st_geometry(red_sea) <- st_geometry(red_sear_tmp)

gulf_aden <- st_read(file.path(mazu_spatial, "data/gulf_aden/gulf_aden.shp"))
plot(st_geometry(gulf_aden), add=TRUE)

eez <- st_read(file.path(mazu_spatial, "data/World_EEZ_v12_20231025/eez_v12.shp"))

Jordan <- grep("Jordan", eez$GEONAME)
Jordan <- eez[Jordan, ]
#plot(st_geometry(Jordan), add=TRUE, border="green")
Sudan <- grep("Sudan", eez$TERRITORY1)
Sudan <- eez[Sudan, ]
#plot(st_geometry(Sudan), add=TRUE, col="green")
Halaib <- grep("Halaib Triangle", eez$GEONAME) #disputed area
Halaib <- eez[Halaib, ]
#plot(st_geometry(Halaib), add=TRUE, col="orange")
Eritrea <- grep("Eritrea", eez$TERRITORY1)
Eritrea <- eez[Eritrea, ]
#plot(st_geometry(Eritrea), add=TRUE, col="green")
Doumeira <- grep("Doumeira Islands", eez$GEONAME) #disputed area
Doumeira <- eez[Doumeira, ]
#plot(st_geometry(Doumeira), add=TRUE, border = "orange", col="orange") # border of Red Sea and Gulf of Aden
Djibouti_tmp <- grep("Djibouti", eez$TERRITORY1)
Djibouti_tmp <- eez[Djibouti_tmp, ]
Djibouti_rs_tmp <- st_intersection(st_geometry(Djibouti_tmp), st_geometry(red_sea))
Djibouti_rs_tmp <- st_sf(Djibouti_rs_tmp)
Djibouti_rs_tmp <- Djibouti_rs_tmp[st_geometry_type(Djibouti_rs_tmp) == "POLYGON" | st_geometry_type(Djibouti_rs_tmp) == "MULTIPOLYGON", ]
Djibouti_rs <- Djibouti_tmp
st_geometry(Djibouti_rs) <- st_geometry(Djibouti_rs_tmp)
plot(st_geometry(Djibouti_rs), border = "orange", col="orange", add=TRUE)

Djibouti_ga_tmp <- st_intersection(st_geometry(Djibouti_tmp), st_geometry(gulf_aden))
Djibouti_ga_tmp <- st_sf(Djibouti_ga_tmp)
Djibouti_ga_tmp <- Djibouti_ga_tmp[st_geometry_type(Djibouti_ga_tmp) == "POLYGON" | st_geometry_type(Djibouti_ga_tmp) == "MULTIPOLYGON", ]
Djibouti_ga <- Djibouti_tmp
st_geometry(Djibouti_ga) <- st_geometry(Djibouti_ga_tmp)
plot(st_geometry(Djibouti_ga), border = "red", col="red", add=TRUE)

Yemen_tmp <- grep("Yemen", eez$GEONAME)
Yemen_tmp <- eez[Yemen_tmp, ]
Yemen_rs_tmp <- st_intersection(st_geometry(Yemen_tmp), st_geometry(red_sea))
Yemen_rs_tmp <- st_sf(Yemen_rs_tmp)
Yemen_rs_tmp <- Yemen_rs_tmp[st_geometry_type(Yemen_rs_tmp) == "POLYGON" | st_geometry_type(Yemen_rs_tmp) == "MULTIPOLYGON", ]
Yemen_rs <- Yemen_tmp
st_geometry(Yemen_rs) <- st_geometry(Yemen_rs_tmp)

Yemen_ga_tmp <- st_intersection(st_geometry(Yemen_tmp), st_geometry(gulf_aden))
Yemen_ga_tmp <- st_sf(Yemen_ga_tmp)
Yemen_ga_tmp <- Yemen_ga_tmp[st_geometry_type(Yemen_ga_tmp) == "POLYGON" | st_geometry_type(Yemen_ga_tmp) == "MULTIPOLYGON", ]
Yemen_ga <- Yemen_tmp
st_geometry(Yemen_ga) <- st_geometry(Yemen_ga_tmp)


Saudi_Arabia <- Saudi_Arabia_tmp[1, ]


#plot(st_geometry(Yemen), add=TRUE, border = "orange", col="orange")

Egypt_pre <- grep("Egyptian Exclusive Economic Zone", eez$GEONAME)
Egypt_pre <- eez[Egypt_pre, ]
Egypt_tmp <- st_intersection(st_geometry(Egypt_pre), st_geometry(red_sea))
Egypt_tmp <- st_sf(Egypt_tmp)
Egypt_tmp <- Egypt_tmp[st_geometry_type(Egypt_tmp) == "POLYGON" | st_geometry_type(Egypt_tmp) == "MULTIPOLYGON", ]
Egypt <- Egypt_pre
st_geometry(Egypt) <- st_geometry(Egypt_tmp)
#plot(st_geometry(Egypt), border = "green", col="green")

Israel_pre <- grep("Israel", eez$GEONAME)
Israel_pre <- eez[Israel_pre, ]
Israel_tmp <- st_intersection(st_geometry(Israel_pre), st_geometry(red_sea))
Israel_tmp <- st_sf(Israel_tmp)
Israel_tmp <- Israel_tmp[st_geometry_type(Israel_tmp) == "POLYGON" | st_geometry_type(Israel_tmp) == "MULTIPOLYGON", ]
Israel <- Israel_pre
st_geometry(Israel) <- st_geometry(Israel_tmp)

plot(st_geometry(Israel), border = "blue", col="blue")


Saudi_Arabia_tmp <- grep("Saudi Arabian Exclusive Economic Zone", eez$GEONAME)
Saudi_Arabia_tmp <- eez[Saudi_Arabia_tmp, ]
Saudi_Arabia_tmp <- st_cast(Saudi_Arabia_tmp, "POLYGON")
Saudi_Arabia <- Saudi_Arabia_tmp[1, ]
plot(st_geometry(Saudi_Arabia), add=TRUE, border = "purple", col="purple")


Somalia_pre <- grep("Somalia", eez$TERRITORY1)
Somalia_pre <- eez[Somalia_pre, ]
Somalia_tmp <- st_intersection(st_geometry(Somalia_pre), st_geometry(gulf_aden))
Somalia_tmp <- st_sf(Somalia_tmp)
Somalia_tmp <- Somalia_tmp[st_geometry_type(Somalia_tmp) == "POLYGON" | st_geometry_type(Somalia_tmp) == "MULTIPOLYGON", ]
Somalia <- Somalia_pre
st_geometry(Somalia) <- st_geometry(Somalia_tmp)

plot(st_geometry(Somalia), border = "purple", col="purple")

eezs <- rbind(Jordan, Sudan, Halaib, Eritrea, Doumeira, Djibouti_rs, Djibouti_ga, Yemen_rs, Yemen_ga,
Egypt, Israel, Saudi_Arabia, Somalia)

eezs <- eezs %>%
  select(description = GEONAME,
         country = TERRITORY1,
         iso3c = ISO_TER1) %>%
  mutate(description = gsub("Exclusive Economic Zone", "EEZ", description)) %>%
  mutate(description = gsub("Overlapping claim", "", description)) %>%
  mutate(iso3c = ifelse(country == "Halaib Triangle", "SDN/EGY", iso3c)) %>%
  mutate(iso3c = ifelse (country == "Doumeira Islands",
         "ERI/DJI", iso3c))  %>%
    mutate(country = ifelse(country == "Halaib Triangle", "Sudan / Egypt", country)) %>%
  mutate(country = ifelse (country == "Doumeira Islands",
         "Eritrea / Djibouti", country))  %>%
  mutate(water_body = c("Red Sea", "Red Sea", "Red Sea", "Red Sea", "Red Sea", "Red Sea", "Gulf of Aden", "Red Sea", "Gulf of Aden", "Red Sea", "Red Sea", "Red Sea", "Gulf of Aden"))

#correct invalid geometries
valid <- st_is_valid(eezs)

# If there are invalid geometries, try to fix them
if (any(!valid)) {
  eezs$geometry <- st_make_valid(eezs$geometry)
}


eezs$area_km2 <- as.numeric(round(st_area(eezs)/1000000, 0))

plot(st_geometry(eezs))
st_write(eezs, here("_spatial/data/eez.shp"))


plot(st_geometry(red_sea), border="red")
plot(st_geometry(gulf_aden), border="blue", add=TRUE)

wb <- rbind(red_sea, gulf_aden)
wb <- wb %>%
  select(name)
wb$area_km2 <- as.numeric(round(st_area(wb)/1000000, 0))
  
st_write(wb, here("_spatial/data/water_bodies.shp"))
```
